---
title: "xts"
output: html_notebook
---

# Time Series

A time series object

* R stores times series in objects. 
* It must have data and a time index.

Some classes of time series objects

* ts
* zoo
* xts

We are going to start with data sets from the *R Datasets Package*.

Load the dataset promise.
```{r}
data("AirPassengers")
```

Look at the structure of the data set.

```{r}
str(AirPassengers)
```
What type of observations does the time series contain?
```{r}
typeof(AirPassengers)
```
What is the class of the series?
```{r}
class(AirPassengers)
```
Use the plot method for ts objects.
```{r}
plot(AirPassengers, main = "Air Line Passenger Data")
```
Load the xts Library.  This will also load the zoo package. xts extends the zoo package.
This allow you to use zoo methods with xts.
```{r}
library(xts)
```
Look at the first few observations.
```{r}
head(AirPassengers)
```
convert to an xts object.
```{r}
AirPassengers_xts <- as.xts(AirPassengers)
```
Look at the first few observations.
```{r}
head(AirPassengers_xts)
```

xts allows standard matrix opperations.
```{r}
AirPassengers_xts[3,1]
```

```{r}
length(AirPassengers_xts)
```
How to create a sequence in R
```{r}
seq(2, 144, by = 12)
```
```{r}
AirPassengers_xts[seq(2,144, by = 12),]

```


Exercise.

* Extract observation 12.
* Extact observation 24.
* Extract one observation for Feb every year.

The index of an xts object is a time object.  Here is how to create 
a time object using seq.
```{r}
my_date_vector <- seq(as.Date('1952-01-01'), length = 150, by = 'month')

```
```{r}
my_date_vector
```

You can create white noise
```{r}
my_white_noise <- rnorm(150)
```
```{r}
sd(my_white_noise)
```


You can create a date
```{r}
independence_day_1960 <- as.POSIXct("1960-07-04")
```

```{r}
 white_noise <- xts(x = my_white_noise, order.by = my_date_vector,
                     i_day = independence_day_1960)
```
```{r}
head(white_noise)
```

```{r}
plot(white_noise)
```
Attributes store information about an object.

```{r}
attributes(white_noise)
```





The xts object contains data and an index.  The index is the date part.
```{r}
index(AirPassengers)
```

```{r}
index(AirPassengers_xts)
```
The data part
```{r}
coredata(AirPassengers_xts)
```
Exercise

Exercise
* Extract the data and index from your xts object.
```{r}
data("austres")
```
```{r}
class(austres)
```
```{r}
austres_xts <- as.xts(austres)
```
```{r}
class(austres_xts)
```
```{r}
index(austres_xts)
```
```{r}
length(austres_xts)
```
```{r}
my_date_vector <- seq(as.Date('1971-01-01'), length = 89, by = 'quarter')

```
```{r}
my_date_vector
```

```{r}
xts(x = rnorm(89), order.by = my_date_vector )
```


```{r}
coredata(austres_xts)
```






* Create an xts object with the same dates as AirPassengers but with a white noise column. Add a attribute of an important date for you.


Here is another data set.
```{r}
data("airquality")
```
```{r}
str(airquality)
```
```{r}
coredata(airquality)
```
```{r}
index(airquality)
```

```{r}
airquuality_xts <- as.xts(airquality)
```

```{r}
index(airquality)
```

Exercise:
* Look at airquality. Is it possible to reconsruct that as an xts object?  Search *The R Data Sets Package* to get the starting date and other needed information.

```{r}
length(airquality[,1])
```
```{r}
my_date_vector <- seq(as.Date('1973-05-01'), length = 153, by = 'days')

```
```{r}
coredata(airquality)
```

```{r}
airquality_xts <- xts(x = airquality, order.by = my_date_vector)
```
```{r}
str(airquality_xts)
```
```{r}
head(airquality_xts)
```
```{r}
plot(airquality_xts[,c('Wind','Temp')])
```

```{r}
plot(airquality_xts[,2:3])
```



Extract pieces of a time series.
```{r}
head(AirPassengers_xts)
```

```{r}
my_date_index <- seq(as.Date('1954-01-01'), length = 15, by = 'month')
my_date_index
```
```{r}
AirPassengers_xts[my_date_index]
```
Exercise

* Create a date vector and use that to extract rows from the AirPassengers time series.
```{r}
1:3
```

```{r}
as.Date('1950-04-01') + 1:3
```
Exercise
*  Use as.Date to create an index and extract observations from Feb 1955 to June 1956.
*  Can you use  `as.matrix` to convert and xts to a matrix?  What happens to the dates?
```{r}
head(AirPassengers_xts)
```
```{r}
#as.matrix(AirPassengers_xts)
```


Read and write to disk.
```{r}
write.zoo(AirPassengers_xts, file = file.path(".","AirPassengers.csv"), sep = ',')
```

```{r}
read.zoo(file = file.path(".","AirPassengers.csv"), sep = ',', FUN = as.yearmon)
```
```{r}
data("lynx")
```


Exercise
*  Go to the R data page select a time series, convert it to xts, write it, then read it in.

# Selecting entries using dates

```{r}
head(x)
AirPassengers_xts['1949']
```
```{r}
AirPassengers_xts['1949-03']
```
```{r}
AirPassengers_xts['1950/1960']
```
```{r}
data("AirPassengers")
AirPassengers_xts <- as.xts(AirPassengers)
```

Exercise

* Select three dates,  three years and three month to month ranges and two year to year ranges from and xts data set.  

* Select the max data point from 1955.
```{r}
AirPassengers_xts['1955']
```
```{r}
AirPassengers_xts['1955-07']
```



* Replace all the values for the summer of 1960 with 0.
```{r}
head(AirPassengers_xts)
```
```{r}
tail(AirPassengers_xts)
```


```{r}
first(last(AirPassengers_xts, "1 year"), "4 months")
```
```{r}
first(AirPassengers_xts, "3 years")
```
```{r}
head(airquality_xts)
```


The last three days of the first three years.
```{r}
last(first(AirPassengers_xts, '3 years'), "3 days")
```

Exercise

* Find the last four days.
* Find the first three weeks.
* Find the first four years.
* Find the last two days of the first four years.



Exercise

* Go to the R Datasets Package and locate 5 interesting time series.  Load them into your R session.
* Write each series to your computer
* Read each series back into R.



```{r}
data("airmiles")
```

```{r}
str(airmiles)
```
```{r}
airmiles_xts <- as.xts(airmiles)
```
```{r}
str(airmiles_xts)
```

```{r}
airmiles_index <- index(airmiles_xts)
head(airmiles_index)
```

```{r}
AirPassengers_xts[airmiles_index]
```


Exercise:
* Use the index of each series to find out which of your series cover the same time periods.
```{r}
xts::merge.xts()
```

```{r}
merge(AirPassengers_xts, airmiles_xts, join = "inner")
```


```{r}
#index(airquality_xts)
```
```{r}
#index(AirPassengers_xts)
```


```{r}
#merge(airquality_xts, AirPassengers_xts, fill = 'locf', drop = TRUE)
```

```{r}
rbind(AirPassengers_xts, airquality_xts)
```

```{r}
head(airquality_xts)
```

```{r}
head(na.locf(airquality_xts, fromLast = FALSE ))
```
```{r}
head(na.locf(airquality_xts, fromLast = TRUE ))
```

Change in air quality

```{r}
airquality_lead <- lag(airquality_xts, k = -1)
head(airquality_lead)
```
```{r}
airquality_lag <- lag(airquality_xts, k = 1)
head(airquality_lag)
```


```{r}
head(airquality_xts)
```

```{r}
aq <- cbind(airquality_lead,  airquality_xts, airquality_lag )
```
```{r}
head(aq)
```
```{r}
head(lag(airquality_xts, k = 3))
```

```{r}
airquality_diff <- diff(airquality_xts, lag = 1, differences = 2)
```
```{r}
head(airquality_diff)
```


```{r}
airquality_w_diff <- merge(airquality_xts, airquality_diff)
```
```{r}
names(airquality_w_diff )
```

```{r}
plot(airquality_w_diff[,c('Ozone','Ozone.1')])
```





```{r}
air_weeks <- endpoints(airquality_xts, on = 'weeks')

```
```{r}
air_weeks
```
```{r}
#airquality_xts['1973-06/1973-08',"Ozone"]
```
```{r}
airquality_mean <- period.apply(airquality_xts[,"Ozone"], INDEX = air_weeks, FUN = mean)
```
```{r}
airquality_mean 
```

```{r}
airquality_mean <- period.apply(airquality_xts[,"Ozone"], INDEX = air_weeks, FUN = mean, na.rm = TRUE)
```
```{r}
class(airquality_mean)
```
```{r}
plot(airquality_mean)
```
```{r}
airquality_mean
airquality_xts[,'Ozone']
air_merge <- merge(airquality_xts[,'Ozone'], airquality_mean)
```
```{r}
head(air_merge)
```

```{r}
plot(air_merge)
```
```{r}
na.approx(air_merge)[1:10,]
```

```{r}
plot(na.approx(air_merge))
```






```{r}
plot(na.locf(air_merge))
```
```{r}
plot(na.locf(air_merge, fromLast = TRUE))
```





