---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(forecast)
library(sweep)
```
Read in the CO2 dataset.  
```{r}
data("co2")
```

```{r}
head(co2)
```
```{r}
head(as.xts(co2))
```
Here is a reminder of the stucture of an xts object
```{r}
co2_xts <- xts(x = co2, order.by = seq(as.Date('1959-01-01'), length = 468, by = 'months'))
```
```{r}
f <- forecast::ets(co2_xts)
```

```{r}
sweep::sw_augment(f)
```
```{r}
sweep::sw_glance(f)
```
```{r}
sweep::sw_tidy(f)
```
```{r}
sweep::sw_tidy_decomp(f)
```
```{r}
forecast::Acf(co2_xts)
```
```{r}
forecast::arfima(co2_xts)
```
```{r}
forecast::auto.arima(co2_xts)
```
```{r}
forecast::checkresiduals(f)
```
```{r}
forecast::findfrequency(co2_xts)
```
```{r}
forecast::ggAcf(co2_xts)
```
```{r}
forecast::ggPacf(co2_xts)
```












```{r}
head(co2_xts)
```
Here is a plot, but notice it is very decpetive. How to lie with data.  Looks like the 
world is going to end soon!
```{r}
plot(co2, ylab = expression("Atmospheric concentration of CO"[2]),
     las = 1)
title(main = "co2 data set")
```
Here is the same plot only with the xts object.
```{r}
plot(co2_xts, ylab = expression("Atmospheric concentration of CO"[2]),
     las = 1)
title(main = "co2 data set")
```
What if we want to use ggplot2?  What would be use for the x value?  
We need tidy data or a `data.frame`.  
We can use  `fortify`

```{r}
head(fortify(co2_xts))
```
How to lie with statistics. 
Looks like the end of the world is near!
```{r}
fortify(co2_xts) %>% ggplot(aes(x = Index, y = co2_xts )) + 
  geom_point(size = 0.1) + 
  geom_line() +
  theme_classic()
```
How not to lie.
What happened to the increase?
```{r}
fortify(co2_xts) %>% ggplot(aes(x = Index, y = co2_xts )) + 
  geom_point(size = 0.1) + 
  geom_line() +
  expand_limits(y = 0) +
  theme_classic()
```
Lets look at one point per year.  Review endpoints.
```{r}
e <- endpoints(co2_xts, on = 'years') + 6
```

Lets plot one point per year.  The rate does not seem to be increasing much.
```{r}
fortify(co2_xts) %>% 
  slice(e) %>% 
  ggplot(aes(x = Index, y = co2_xts )) + 
  geom_point(size = 0.1) + 
  geom_line() +
  expand_limits(y = 0) +
  geom_smooth(method = 'lm') + 
  theme_classic()
```
How fast is CO2 really increasing?  This also shows how to use an `r` value in markdown.
```{r}
p <- (as.numeric(last(co2_xts)) -  as.numeric(first(co2_xts)) ) / as.numeric(first(co2_xts)) * 100 / length((co2_xts))
```
```{r}
start(co2_xts)
```
```{r}
first(co2_xts)
```

```{r}
end(co2_xts)
```
```{r}
last(co2_xts)
```

```{r}
identical(start(co2_xts), first(co2_xts))
```



There is a `r format(p, digits = 2)` % increase per year


```{r}
fortify(co2_xts) %>% ggplot(aes(x = Index, y = co2_xts )) + 
  geom_point(size = 0.1) + 
  geom_line() +
  expand_limits(y = 0) +
  theme_classic()
```
```{r}
head(time(co2_xts))
```
```{r}
head(index(co2_xts))
```
```{r}
identical(time(co2_xts), index(co2_xts) )
```
```{r}
deltat(co2_xts)
```
```{r}
frequency(co2_xts)
```
```{r}
head(cycle(co2_xts))
```

```{r}
data(UKLungDeaths)
```
```{r}
str(fdeaths)
```
```{r}
frequency(fdeaths)
```
```{r}
periodicity(fdeaths)
```


```{r}
str(mdeaths)
```
```{r}
plot(fdeaths)
```
```{r}
start(fdeaths)
```
```{r}
fdeaths_xts <- as.xts(fdeaths)
```
```{r}
plot(fdeaths_xts)
```
```{r}
mdeaths_xts <- as.xts(mdeaths)
```

```{r}
fmdeaths <- merge(mdeaths_xts, fdeaths_xts)
```
```{r}
frequency(fmdeaths)
```
```{r}
periodicity(fmdeaths)
```
```{r}
head(fmdeaths)
```
```{r}
fdeaths
```
```{r}
fdeaths_xts
```
You might want to control  frequency
```{r}
attr(fdeaths_xts, 'frequency') <- 4
```
```{r}
as.ts(fdeaths_xts)
```



Attributes of a time series.

* Mean: fixed or varying?
* variance: constant or changing?
* correlation over time?

Things to learn
* Arima.sim
* arima


The index dates only show the year and the month.  This is done by applying the
`as.yearmonth` function to a set of dates.  
```{r}
class(index(fmdeaths))
```


```{r}
plot(fmdeaths, legend.loc = 'topright')
```
```{r}
frequency(fmdeaths)
```





Go to the R Datasets Package page.
Select ten time series and plot them.
Use an R notebook.
Organize the ten plots by some characteristic of the plot or data.
You will explain this to the class.
 
 Possible grouping factors.
* periodicity
* Similar time periods
* Similar plot patterns
    - increasing
    - decreasing
    - increasing variance
    - exponential (non-linear) growth
    - seasonality
* predictability 


To review
* `log` tranformation for series with exponential growth or increasing variance.
* `diff` transformation for series with a linear trend.
* seasonal `diff`

Lets try differencing with the co2 data.
We can us seasonal differencing.

```{r}
white_noise <- arima.sim(model = list(order = c(0,0,0)), n = 200, mean = 100, sd = 1000)
```
```{r}
plot(white_noise)
```
```{r}
not_white_n <- arima.sim(n= 200, list(ar = c(.2), ma = c(.1)), sd = sqrt(100))
```
```{r}
plot(not_white_n)
```




Non-Stationary.  The mean might change.  (Different parts of the series follow different models.)

Random walk with a drift
$Y_t = c + Y_{t-1} + \epsilon_t$

$\epsilon_t \sim N(0,\sigma^2) $


Use cumsum() to get a random walk

cbind to merge time series together


$Y_t - \mu = \phi(Y_{t-1} - \mu) + \epsilon_t $

White Noise

$Y_t - \mu = \epsilon_t $

Random Walk


$Y_t = Y_{t-1}  + \epsilon_t $

Random walk with drift
$Y_t - \mu = (Y_{t-1} - \mu) + \epsilon_t $

