---
title: "R Notebook"
output: html_notebook
---

# Read in a data set with tidyquant.
* Show tq_get for stock prices.
* Show integration with dplyr and the tidyverse.
* Read data from FRED economic data.

Tidyquant is a package that tries to integrate financial analysis and therefore time series 
analysis into the tidyverse. We will get a taste of that today.


```{r}
library(tidyquant)
library(forecast)
library(timetk)
#install.packages("timetk")
```
The function tq_get receives an identifier and produces the indicated data.
Here it received the identifier for the stock Netflix and returned the stock prices for 
Netflix as indicated in the function call.
```{r}
tq_get("NFLX", get = "stock.prices")
```
The function is vectorized so it can receive more than one identifier
```{r}
tq_get(c("NFLX","AAPL"), get = "stock.prices")
```
The concept of vectorization can be expanded by using a function that create a list 
if identifiers for calling tq_get.
Here we create a vector of all indexes in the SP500.  This vector will then 
be used to call tq_get.

```{r}
SP500_index_labels <- tq_index("SP500")
```
Notice that there is a bunch of information for each SP500 index.

```{r}
SP500_index_labels
```
```{r}
class(SP500_index_labels)
```
Remember the dplyr slice function
```{r}
slice(SP500_index_labels, 1:3)
```
With dplr we can send this list of indexes to tq_get to return a list is prices.
```{r}
stock_data <- SP500_index_labels %>% 
  slice(1:3) %>% 
  tq_get(get = "stock.prices")
```

```{r}
stock_data
```
```{r}
class(stock_data)
```


Tidyquant is an extension to other time series related packages in R.  
It tries to apply wrappers that allow them to be used in the tidyverse.
It applies wrappers around functions available in other packages
so they can be used with tibbles, which are the tidy verse data frames.
With tq_transmute_fun_options, we can see all the functions that have wrappers.
This also gives a list of important function for each of these packages.
Notice there are not any functions from the forecast package listed. 
The forcast package is the one with the model fitting functions.
It appears you will still have to convert to a ts or xts object to 
do ARIMA modeling, but this package and make data manipulation of 
time series tibles follow the tidy data principle.
```{r}
tq_transmute_fun_options()
```
Remember in the table stock_data there is the field adjusted.
glimpse is a tidyverse function to view any tidy verse object quickly
```{r}
glimpse(stock_data)
```
Returns is a financial term for profit on investment.

```{r}
monthly_returns <- stock_data %>% 
  group_by(symbol) %>% 
  tq_transmute(adjusted, mutate_fun = monthlyReturn)
```

```{r}
monthly_returns
```

```{r}
monthly_returns %>% 
  ggplot(aes(x=date, y = monthly.returns)) +
  geom_point()
```
Get some economic data.
```{r}
oil <- tq_get("DCOILWTICO", get = "economic.data")
names(oil)
dim(oil)
str(oil)
glimpse(oil)
head(oil)
oil %>% na.omit() %>% Pacf


oil %>% na.omit() %>% select(price) %>% pacf


oil %>% na.omit() %>% select(price) %>% Acf

plot(oil)

plot(as.xts(oil))
head(timetk::tk_xts(na.omit(oil)))


oil_xts <- timetk::tk_xts(na.omit(oil))

is.xts(oil_xts)

pacf(oil_xts)
acf(oil_xts)

```

```{r}
forecast::auto.arima(oil_xts)
```


```{r}
gold <- tq_get("GOLDAMGBD228NLBM", get = "economic.data")
glimpse(gold)
```
```{r}
gold_xts <- timetk::tk_xts(na.omit(gold))
glimpse((gold_xts))
```
```{r}
Acf(gold_xts)
```
```{r}
Pacf(gold_xts)
```
```{r}
plot(gold_xts)
```
```{r}
auto.arima(gold_xts)
```
```{r}
gold_diff <- diff(gold_xts)
```
```{r}
Pacf(gold_diff)
```
```{r}
Acf(gold_diff)
```


```{r}
clf <- tq_get("LNU01000000", get = "economic.data")
```
```{r}
glimpse(clf)
```
```{r}
clf_xts <- timetk::tk_xts(na.omit(clf))
```
```{r}
glimpse(clf_xts)
```


```{r}
forecast::auto.arima(clf_xts)
```




```{r}
rmhi <- tq_get("MEHOINUSA672N", get = "economic.data")
```
```{r}
rmhi_xts <- timetk::tk_xts(na.omit(rmhi))
```

```{r}
glimpse(rmhi)
```
```{r}
rmhi_xts["2008-01/2016-01"]
```
```{r}
forecast::ets(rmhi_xts)
```






plot(gold)


gold %>% na.omit() %>% pacf
```

```{r}
monthly_returns %>% na.omit() %>% pacf
```

```{r}
mr <- monthly_returns %>% 
  select(date, monthly.returns) 

mr_xts <- timetk::tk_xts(mr, date_var = date, freq = 12)
mr <- as.vector(monthly_returns$monthly.returns)
my_date <- as_date(monthly_returns$date)
mr_xts <- xts(x = mr,
              order.by = my_date,
              frequency = 12)

acf(mr_xts)
monthly_returns %>% 
  select(date, monthly.returns) %>% 
  timetk::tk_xts(date_var = date)

fortify(monthly_returns)
monthly_returns_xts <- as.xts(monthly_returns$monthly.returns, 
                              order.by = monthly_returns$date,
                              frequency = 12)
mr <- monthly_returns[c("date","monthly.returns")]
mr
head(monthly_returns[c("date","monthly.returns")])
```
```{r}
class(mr_xts)
acf(mr_xts)
```

```{r}
acf(monthly_returns_xts)
```

```{r}

```










